---
title: "Nearshore and Deepwater Habitat Use"
author: "Brett Johnson"
date: '`r date()`'
output:
 html_document:
   theme: cosmo
   code_folding: show
   toc: true
   toc_float: true
   number_sections: true
---
# Setup

```{r setup, echo FALSE}
# read in data files
library(hakaisalmon)
library(tidyverse)
library(lubridate)
library(here)
library(hakaiApi)
library(DescTools)
library(car)

options(scipen = 999)

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, paged.print=TRUE)
theme_set(ggsidekick::theme_sleek(base_size = 18)) #Thanks Sean Anderson


spp_labels <- c(CU = "Chum", PI = "Pink", SO = "Sockeye", DI = "Discovery Islands", 
                JS = "Johnstone Strait", "dfo" = "Deepwater", "hakai" = "Nearshore")

okisollo <- read_csv(here("processed_data", "okisollo.csv"))
okisollo$Habitat <- dplyr::recode(okisollo$Habitat, "Mid Channel" = "Deepwater" )

dfo_coords <- read_csv(here("processed_data", "dfo_coords"))
seines <- read_csv(here("processed_data", "seines.csv"))

dfo_data <- okisollo %>% 
  filter(Habitat == "Deepwater")
```

# Map

```{r map}
library(maps) 
library(mapdata)
library(maptools)
library(rgeos)
library(rgdal)
library(ggplot2)
library(ggsn)
library(dplyr)
library(raster)
library(here)

hakai_coords <- okisollo %>% 
  dplyr::filter(Habitat == "Nearshore") %>% 
  mutate(org = "Nearshore") %>% 
  # Remove 'nearshore' seines that were actually overlapping with mid-channel sites:
 dplyr::filter(!seine_id %in% c("DE231N2", "DE151N1", "DE301N1", "DE290N2", "DE236N1", "DE361N2", "DE220N1")) %>% 
  dplyr::select(seine_id, org, lon = lat, lat = lon) %>% 
  dplyr::filter(lat > 50.3, lon < -125.315)


#TODO: Once Nate fixes the lat/lon reversal in the portal I need to fix the above select statement
both_coords <- rbind(hakai_coords, dfo_coords) %>% 
  distinct()

mean_hakai_coords <- c(mean(hakai_coords$lat), mean(hakai_coords$lon))
mean_dfo_coords <- c(mean(dfo_coords$lat), mean(dfo_coords$lon))

# Download shape file and plot map


BC.shp <- readOGR(here("data","2_Shapefile", "COAST_TEST2.shp"))

quadra <- extent(-126.15, -124.650, 49.96, 50.46)
quadra_shp <- crop(BC.shp, quadra)
quadra_df <- fortify(quadra_shp)

(okisollo_map <- ggplot() + theme_bw() +
  geom_polygon(data = quadra_df, aes(long, lat, group = group),
               colour = "black", size = 0.1, fill = 'grey95') +
  coord_map(xlim = c(-125.375, -125.275), ylim = c(50.28, 50.32))+
  geom_point(data = both_coords, aes(x = lon, y = lat, colour = org), 
              size = 4, alpha = 0.5) +
  scalebar(x.min = -125.28, x.max = -125.325, 
                y.min = 50.285, y.max = 50.29, 
                dist = 1, dist_unit="km", transform = TRUE,
                model = "WGS84", height = 0.2, 
                st.dist = 0.3,
                box.fill = c("grey30", "white"),
                st.color = "grey30",
                border.size = 0.5,
                st.size = 3.5) +
    theme(legend.title = element_blank())
)

okisollo_map <- okisollo_map + geom_text() + annotate("text", label = "Okisollo Channel", 
                                      x = -125.3, y = 50.307, size = 4,
                                      colour = "black") +
  annotate("text", label = "Quadra Island", x = -125.325, y = 50.2818, size = 4,
           colour = "black")+
  annotate("text", label = "Sonora Island", x = -125.315, y = 50.315, size = 4,
           colour = "black") +
  annotate("text", label = "Venture Point", x = -125.348, y = 50.2954, size = 2.5,
           colour = "black")

north2(okisollo_map, x = 0.15, y = 0.73, symbol = 3) 
ggsave(here("final_figs", "map_inset.png"))

detach("package:raster", unload=TRUE)
beepr::beep(2)

```

# Condition

```{r condition}

# Compare annual average condition for each habitat and each species with 95 % CIs
library(dplyr)
condition <- okisollo %>% 
  dplyr::filter(species %in% c("SO", "CU")) %>%
  mutate(year = year(survey_date)) %>% 
  dplyr::filter(!(year == 2016 & species %in% c("PI", "CO"))) %>% 
  group_by(year, Habitat, species) %>% 
  summarise(mean = mean(k, na.rm = T),
            sd = sd(k, na.rm = T), 
            n = n()) %>% 
  mutate(se = 1.96 * (sd/sqrt(n))) %>% 
  mutate(CI = round(qt(1 - (0.05 / 2), n - 1) * se,2))

okisollo %>%
  dplyr::filter(species %in% c("SO", "CU")) %>%
  mutate(year = year(survey_date)) %>% 
  dplyr::filter(!(year == 2016 & species %in% c("PI", "CO"))) %>% 
  ggplot(aes(x = species, y = k, fill = Habitat)) +
  geom_boxplot() +
  ylab("Condition factor") +
  xlab("Species")+
  facet_grid(~year, scales = "free") +
  scale_x_discrete(breaks=c("CO", "CU", "PI", "SO"),
          labels=c("Coho", "Chum", "Pink", "Sockeye")) +
  scale_fill_hakai() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.direction = "vertical", legend.position = c(.8,.85)) 

ggsave(here("final_figs", "condition.png"),  width = 5, height = 5)

# sockeye in 2015 comparison of habitats
(cond_stats <- okisollo %>% 
  dplyr::filter(year == 2015,
                species %in% c("SO")) %>% 
    ungroup() %>% 
  do(broom::tidy(t.test(.$k ~ .$Habitat))))

# sockeye in 2015 comparison of habitats
(cond_stats <- okisollo %>% 
  dplyr::filter(year == 2016,
                species %in% c("CU","SO")) %>% 
    ungroup() %>% 
    group_by(species) %>% 
  do(broom::tidy(t.test(.$k ~ .$Habitat))))

```


# Lengths

```{r annual species lengths plot}

# Compare annual average lengths for each habitat and each species with 95 % CIs
library(dplyr)
lengths <- okisollo %>% 
  dplyr::filter(fork_length < 300, species %in% c("SO", "PI", "CU", "CO")) %>%
  mutate(year = year(survey_date)) %>% 
  dplyr::filter(!(year == 2016 & species %in% c("PI", "CO"))) %>% 
  group_by(year,Habitat, species) %>% 
  summarise(mean = mean(fork_length, na.rm = T),
            sd = sd(fork_length, na.rm = T), 
            n = n()) %>% 
  mutate(se = 1.96 * (sd/sqrt(n))) %>% 
  mutate(CI = round(qt(1 - (0.05 / 2), n - 1) * se,2))

okisollo %>%
  dplyr::filter(fork_length < 300, species %in% c("SO", "PI", "CU", "CO")) %>%
  mutate(year = year(survey_date)) %>% 
  dplyr::filter(!(year == 2016 & species %in% c("PI", "CO"))) %>% 
  
  ggplot(aes(x = species, y = fork_length, fill = Habitat)) +
  geom_boxplot() +
    # geom_errorbar(position = position_dodge(.9), width = .25, aes(ymin = mean - CI, ymax = mean +
    #                     CI)) +
  ylab("Fork length (mm)") +
  xlab("Species")+
  facet_grid(~year, scales = "free") +
  scale_x_discrete(breaks=c("CO", "CU", "PI", "SO"),
          labels=c("Coho", "Chum", "Pink", "Sockeye")) +
  scale_fill_hakai() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.direction = "vertical", legend.position = c(.8,.85))

#TODO: Put the legend in the top left of the plot
#TODO: Add sample sizes to plot

ggsave(here("final_figs", "lengths_annual.png"))

# knitr::kable(lengths, style = "markdown")
```


```{r annual species lengths stats}

(len_stats_2015 <- okisollo %>% 
  dplyr::filter(year == 2015) %>% 
  #mutate(Habitat = factor(Habitat)) %>% 
  group_by(species) %>% 
  do(broom::tidy(t.test(.$fork_length ~ .$Habitat))))

(len_stats_2016 <- okisollo %>% 
  dplyr::filter(year == 2016, species %in% c("SO", "CU")) %>% 
  mutate(Habitat = factor(Habitat)) %>% 
  group_by(species) %>% 
  do(broom::tidy(t.test(.$fork_length ~ .$Habitat))))
```


# Sockeye Stock Composition

```{r annual gsi}

gsi_okisollo <- okisollo %>%
  dplyr::filter(collapsed_fraser_stocks != 0, collapsed_fraser_stocks != "") %>% 
  drop_na(site_id) %>% 
  drop_na(collapsed_fraser_stocks) %>%
  group_by(year, Habitat, collapsed_fraser_stocks) %>% 
  summarize(n = n()) %>% 
  mutate(proportion = n / sum(n)) %>% 
  arrange(proportion) 

# Breakout out habitat and year proportions to calculate CIs
## nearshore 2015
gsi_okisollo_nearshore_2015 <- gsi_okisollo %>% 
  dplyr::filter(year == 2015, Habitat == "Nearshore")
h_15 <- as_tibble(MultinomCI(gsi_okisollo_nearshore_2015$n, 
           conf.level=0.95, 
           method="sisonglaz")) 
gsi_okisollo_nearshore_2015 <- bind_cols(gsi_okisollo_nearshore_2015, h_15)

## nearshore 2016
gsi_okisollo_nearshore_2016 <- gsi_okisollo %>% 
  dplyr::filter(year == 2016, Habitat == "Nearshore")
h_16 <- as_tibble(MultinomCI(gsi_okisollo_nearshore_2016$n, 
           conf.level=0.95, 
           method="sisonglaz")) 
gsi_okisollo_nearshore_2016 <- bind_cols(gsi_okisollo_nearshore_2016, h_16)

## mid-channel 2015
gsi_okisollo_deepwater_2015 <- gsi_okisollo %>% 
  dplyr::filter(year == 2015, Habitat == "Deepwater")
m_15 <- as_tibble(MultinomCI(gsi_okisollo_deepwater_2015$n, 
           conf.level=0.95, 
           method="sisonglaz")) 
gsi_okisollo_deepwater_2015 <- bind_cols(gsi_okisollo_deepwater_2015, m_15)


## mid-channel 2016
gsi_okisollo_deepwater_2016 <- gsi_okisollo %>% 
  dplyr::filter(year == 2016, Habitat == "Deepwater")
m_16 <- as_data_frame(MultinomCI(gsi_okisollo_deepwater_2016$n, 
           conf.level=0.95, 
           method="sisonglaz")) 
gsi_okisollo_deepwater_2016 <- bind_cols(gsi_okisollo_deepwater_2016, m_16)

gsi_okisollo <- bind_rows(gsi_okisollo_deepwater_2016,
                          gsi_okisollo_deepwater_2015, 
                          gsi_okisollo_nearshore_2016, 
                          gsi_okisollo_nearshore_2015) %>% 
  dplyr::filter(collapsed_fraser_stocks %in% c('Late_Shuswap_Portage', 'Chilko',
                                        'Quesnel', 'Early_Thompson',
                                        'Late_Stuart_Stellako'))

rm(gsi_okisollo_deepwater_2016,
                          gsi_okisollo_deepwater_2015, 
                          gsi_okisollo_nearshore_2016, 
                          gsi_okisollo_nearshore_2015, m_16, m_15, h_15)
  
ggplot(gsi_okisollo, aes(x = fct_reorder(collapsed_fraser_stocks, proportion), y = proportion, fill = Habitat)) +
  geom_bar(stat="identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = lwr.ci, ymax = upr.ci), width = 0.5, position = position_dodge(0.9)) +
  ggtitle("Genetic Stock Proportions") +
  facet_wrap(~year) +
  xlab("") + 
  ylab("Proportion") +
  coord_flip() +
  theme(legend.position="bottom") + 
  scale_fill_hakai()

ggsave(here("final_figs", "gsi_props.png"))

```

Chilko sockeye in higher proportion Deepwater not significantly

Late Shuswap Portage consistently higher proportion in nearshore

## Stock specific length comparisons

So why aren't we seeing LSP Stock as much in deep water habitat?

```{r LSP lengths}

lsp_lengths <- okisollo %>% 
  mutate(Stock = ifelse(collapsed_fraser_stocks == "Late_Shuswap_Portage", "Late_Shuswap_Portage", 
                      "All Other Stocks"))  %>% 
  dplyr::filter(Habitat == "Deepwater") %>% 
  drop_na(Stock)

  ggplot(lsp_lengths, aes(x = factor(year), y = fork_length, fill = Stock)) +
  geom_boxplot() +
  scale_fill_hakai() +
  xlab("Year") +
  ylab("Mean fork length (mm)")+
  ggtitle("Deepwater Habitat Fork Lengths") +
  theme(legend.direction = "vertical", legend.position = c(.6,.9))

ggsave(here("final_figs", "Lengths_lsp_deepwater.png"), width = 5, height = 5)

(lsp_len_stats_2015 <- lsp_lengths %>% 
    group_by(year) %>% 
  do(broom::tidy(t.test(.$fork_length ~ .$Stock))))
         
```

Late Shuswap Portage Stocks are shorter than the rest in the mid-channel habitat which might explain why we see less of them mid-channel and more of them nearshore.

# Migration Timing

```{r}

#build dataset of tidy catches per seine per species

dfo_2015_tidy_catch <- read_csv(here::here("data", "dfo_2015_tidy_catch.csv")) %>% 
  mutate(yday = yday(date)) %>% 
  group_by(species) %>% 
  mutate(org = "dfo", percent = cumsum(n / sum(n) * 100), non_cumul_prop = n / sum(n)) %>% 
  ungroup()
  
dfo_2015_tidy_catch <- dfo_2015_tidy_catch %>% 
  dplyr::select(species, yday, n, org, percent, non_cumul_prop) 

hakai_2015 <- seines %>% 
  filter(site_id == "D09", year(date) == 2015, collection_protocol == "SEMSP") %>%
  select(date, so_total, cu_total) %>% 
  gather(`so_total`, `cu_total`, key = "species", value = "n") %>% 
  drop_na() %>% 
  mutate(species = dplyr::recode(species, "so_total" = "SO", "cu_total" = "CU")) %>% 
  mutate(yday = yday(date)) %>% 
  group_by(species) %>% 
  arrange(yday, species) %>% 
  mutate(org = "hakai", percent = cumsum(n / sum(n) * 100), non_cumul_prop = n / sum(n)) %>% 
  ungroup()

hakai_2015 <- hakai_2015 %>% 
  select(species, yday, n, org, percent, non_cumul_prop)

tidy_okisollo_catch_2015 <- rbind(hakai_2015, dfo_2015_tidy_catch)


# MODEL MIGRATION TIMING

# chum
chum_dfo <- tidy_okisollo_catch_2015 %>% 
  filter(species == "CU", org == "dfo")

chum_dfo_predicted <- log_cumul_abund(chum_dfo$percent, chum_dfo$yday) %>% 
  mutate(org = "dfo", species = "CU") %>% 
  mutate(daily_percent = y - lag(y)) 

#chum hakai
chum_hakai <- tidy_okisollo_catch_2015 %>% 
  filter(species == "CU", org == "hakai")

chum_hakai_predicted <- log_cumul_abund(chum_hakai$percent, chum_hakai$yday) %>% 
  mutate(org = "hakai", species = "CU") %>% 
  mutate(daily_percent = y - lag(y)) 

# sockeye dfo
sockeye_dfo <- tidy_okisollo_catch_2015 %>% 
  filter(species == "SO", org == "dfo")

sockeye_dfo_predicted <- log_cumul_abund(sockeye_dfo$percent, sockeye_dfo$yday) %>% 
  mutate(org = "dfo", species = "SO") %>% 
  mutate(daily_percent = y - lag(y)) 

#sockeye hakai
sockeye_hakai <- tidy_okisollo_catch_2015 %>% 
  filter(species == "SO", org == "hakai")

sockeye_hakai_predicted <- log_cumul_abund(sockeye_hakai$percent, sockeye_hakai$yday) %>% 
  mutate(org = "hakai", species = "SO") %>% 
  mutate(daily_percent = y - lag(y)) 

predicted <- rbind(chum_dfo_predicted, chum_hakai_predicted, sockeye_dfo_predicted, sockeye_hakai_predicted) %>% 
  select(x, y, Habitat = org, species, daily_percent)

predicted$Habitat <- dplyr::recode(predicted$Habitat, hakai = "Nearshore", dfo = "Deepwater")

medians_2 <- tidy_okisollo_catch_2015 %>% 
  select(org, species, yday, n) 

medians_2 <- medians_2[rep(row.names(medians_2), medians_2$n), 1:3] %>% 
  group_by(org, species) %>% 
  summarize(median = median(yday))

tidy_okisollo_catch_2015 <- tidy_okisollo_catch_2015 %>% 
  select(yday, percent, Habitat = org, species)

tidy_okisollo_catch_2015$Habitat <- dplyr::recode(tidy_okisollo_catch_2015$Habitat, hakai = "Nearshore", dfo = "Deepwater")

ggplot() +
  geom_line(data = predicted, aes(x = x, y = y, colour = Habitat)) +
  geom_point(data = tidy_okisollo_catch_2015, aes(x = yday, y = percent, colour = Habitat))+
  facet_wrap(species ~ ., nrow = 2, labeller = labeller(species = spp_labels)) +
  #geom_vline(data = medians_2, aes(xintercept = median, colour = org),
             # size=.75) +
  ylab("Cumulative catch (%)") +
  xlab("Julian Day") +
  scale_colour_hakai() +
  ggtitle("2015 Migration Timing")

ggsave(here("final_figs", "migration_timing.png"))
```

Migration Timing was not different 

# Catch Composition

```{r}
(annual_prop <- okisollo %>% 
  group_by(year, Habitat, species) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n / sum(n)) %>% 
  ggplot(aes(x = Habitat, y = prop, fill = species)) +
    geom_bar(stat="identity")) +
    facet_wrap(. ~ year) + 
  scale_fill_hakai() +
  ylab("Proportion")

ggsave(here("final_figs", "catch_prop.png"))
```


# Diversity

```{r stock diversity}

# Create dataframe format for diversity indices
dfo_diversity <- dfo_data %>% 
  drop_na(stock_1) %>% 
  group_by(stock_1) %>% 
  summarize(n = n()) %>% 
  spread(key = stock_1, value = n) %>% 
  ungroup() %>%
  #TODO: Figure out where these 0 and Xs are comning from
  dplyr::select(- "0", - X)

hakai_diversity <- okisollo %>% 
  drop_na(stock_1) %>% 
  group_by(stock_1) %>% 
  summarize(n = n()) %>% 
  spread(key = stock_1, value = n, fill = 0) %>% 
  ungroup() %>% 
  #TODO: Figure out where these 0 and Xs are comning from
  dplyr::select(- "0", - X)

```


```{r}
library(iNEXT)
# Deepwater
dfo_rarefy <- dfo_diversity %>% 
  gather(key = "stock", value = "Deepwater")

hakai_rarefy <- hakai_diversity %>% 
  gather(key = "stock", value = "Nearshore")

rarefy <- full_join(dfo_rarefy, hakai_rarefy) %>% 
  replace_na(list("Deepwater" = 0, Nearshore = 0)) %>% 
  as.data.frame()

rownames(rarefy) <- rarefy[,1]
rarefy[,1] <- NULL
output_all <- iNEXT(rarefy, q = c(0, 1, 2), datatype = "abundance")

output_rich <- iNEXT(rarefy, q = 0, datatype = "abundance")

# Stock richness
ggiNEXT(output_rich, type = 1) +
  ggsidekick::theme_sleek() +
  scale_colour_hakai() +
  ggtitle("Sockeye Genetic Stock Richness") +
  ylab("Unique stocks") +
  ggsidekick::theme_sleek(base_size = 14)

ggsave(here("final_figs", "stock_accumulations.png"))

# all three hill numbers compared for asymptotic diversity
output_all %>% 
  # extend the deepwater extrapolation to the same length as nearshore
  # change the names of the facet variables to Richness ,shann...
  ggiNEXT(output_all, type=1, facet.var="order")


rarefaction_tbl <- estimateD(rarefy, datatype = "abundance", base = "size",
                             level=NULL, conf = 0.95) 

rarefaction_tbl$order <- dplyr::recode_factor(rarefaction_tbl$order, `0` = "Spp. Richness", `1` = "Shannon", `2` = "Simpson")

rarefaction_tbl <- rarefaction_tbl %>% 
  dplyr::rename(Hill_Number = order, Site = site)

ggplot(rarefaction_tbl, aes(x = Hill_Number, y = qD, fill = Site)) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  geom_errorbar(aes(ymin = qD.LCL, ymax = qD.UCL), width = 0.5, position = position_dodge(.9)) +
  ylab("Diversity") +
  scale_fill_hakai()+
  xlab("Diversity measure") +
  ggtitle("Asymptotic Sockeye Genetic Stock ID Diversity")

ggsave(here("final_figs", "Diversity_measures.png"))
```

```{r end}
beepr::beep(8)
```

